---
title: "newfile2"
author: "Beau Coomer"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Importing Data

You always want to load your packages at the top of your script/ document.

```{r}
#| warning: false

library(tidyverse)
library(readr)
```

You can have functions from R packages (like `dplyr`), but you can also write R functions yourself. Load the custom function [clean_keyence_csv](custom_functions/clean_keyence_csv.R)...

```{r}
# Load clean_keyence_csv()
source("custom_functions/clean_keyence_csv.R") 
```

## Dataset 1: Egg Measures

Use the custom function that was imported like this...

1.  Get your file names

```{r}
file_paths <- list.files(path = "data/raw", pattern = "\\.csv$", full.names = TRUE)
```

2.  Process it all into one tibble (similar to data.frame)

```{r}
  compiled_data <- suppressWarnings(
    file_paths %>%
      set_names(basename(.)) %>%
      map(clean_keyence_csv) %>%
      list_rbind(names_to = "source_file")
  )
```

Then, you can check out your result like this...

```{r}
#| eval: false
View(compiled_data)
```

# Cleaning Data

```{r}
clean_csvs<-compiled_data %>%
  select(!c(Measure,date_saved))
call("clean_csvs")
```

```{r}
str_split_fixed(clean_csvs$source_file,"\\-",4)
```

## Dataset 2: Body Measures

```{r}
body_measures <- read.csv("data/crayfish_body_measurements.csv")
```

# Cleaning Data

First, subset the `compiled_data` object to drop irrelevant columns. The `!` tells `select()` to keep everything *except* the columns listed. 

```{r}
clean_csvs<-compiled_data %>%
  select(!c(Measure,date_saved))

call("clean_csvs") # What is call() for??? -Emmy
```

It's best practice to have all names in R to be easily type-able. We can accomplish this by...

* No punctuation at the beginning or end of a string
* Using all lowercase

You can do a direct rename of columns in R by `rename(new_name = old_name)`. If needing to change a bunch of column names, use `rename_with()`.

```{r}
clean_csvs<-clean_csvs %>%
  rename(egg_no=No.) %>%
  rename_with(tolower)
```

All of the important identifying information is kept in the file names. 

```{r}
# One way to split file names into usable chunks...
str_split_fixed(clean_csvs$source_file,"\\-",4)
```

```{r}
# Better way to split the file names into usable chunks + keep the info
# -- in the data frame
tmp<-clean_csvs %>%
  separate_wider_delim(cols = source_file,delim = "-",names = c("species","treatment","ind_num",NA))
```

```{r}
#| eval: false
View(tmp)
```

Sometimes R will yell at if the data type isn't appropraite. You can check the data type of anything using `class()`.

```{r}
# Works on objects
class(tmp)

# And it can also work on parts of the object
class(tmp$result)
```


If the class isn't what it should be, you can fix it on the fly using one of the many `as.___()` functions. Just know that sometimes, it is not going to work and it may not always be appropriate!


```{r}
tmp<-tmp %>%
  mutate(result=as.numeric(result))

# "$result" is ambiguous and could be misleading, so rename for clarity
tmp<-tmp %>%
rename(egg_diameter_mm=result)
```

```{r}
unique_id <- unite(tmp,treatment,ind_num,col = "treatment_ind_number",sep = "_")
```

```{r}
body_measures <- read.csv("data/crayfish_body_measurements.csv")
```

```{r}
egg_diameter_avgs<-unique_id %>%
egg_diameter_avgs<-tmp %>%
  group_by(treatment_ind_number) %>%
  summarize(avg = mean(egg_diameter_mm))
```

```{r}
ind_egg_counts<-unique_id %>%
  group_by(treatment_ind_number) %>%
  summarize(egg_count = n())
```

```{r}
View(tmp)
View(egg_diameter_avgs)
View(ind_egg_counts)
view(unique_id)
```

```{r}
boxplot(egg_diameter_mm ~ treatment,data = tmp)
```

```{r}
treatment_counts<-ind_egg_counts %>%
  separate_wider_delim(cols = "treatment_ind_number",delim = "_",names = c("treatment","ind_number"))
treatment_sums<-treatment_counts %>%
  group_by(treatment) %>%
  summarize(n = sum(egg_count))
```

```{r}
treatment_avgs<-egg_diameter_avgs %>%
  separate_wider_delim(cols = treatment_ind_number,delim = "_",names = c("treatment","ind_number"))
treatment_avgsv<-treatment_avgs %>%
  group_by(treatment) %>%
  summarize(avg = mean(avg))
```

```{r}
quartiles<-tmp %>%
  select(treatment,egg_diameter_mm) %>%
  group_by(treatment) %>%
  reframe(Q1 = quantile(egg_diameter_mm,prob = 0.25),Median = median(egg_diameter_mm),Q3 = quantile(egg_diameter_mm,prob = 0.75))
view(quartiles)
```



ind_egg_counts
```
