---
title: "newfile2"
author: "Beau Coomer"
format: html
editor: visual
---

You always want to load your packages at the top of your script/ document.
```{r}
library(tidyverse)
```

You can have functions from R packages (like `dplyr`), but you can also write R functions yourself. Load the custom function [clean_keyence_csv](custom_functions/clean_keyence_csv.R)...

```{r}
# Load clean_keyence_csv()
source("custom_functions/clean_keyence_csv.R") 
```

Use the function like this...

1. Get your file names 
  ```{r}
  file_paths <- list.files(path = "data/raw", pattern = "\\.csv$", full.names = TRUE)
  ```

2. Process it all into one tibble (similar to data.frame)

  ```{r}
  compiled_data <- suppressWarnings(
    file_paths %>%
      set_names(basename(.)) %>%
      map(clean_keyence_csv) %>%
      list_rbind(names_to = "source_file")
  )
  ```

Then, you can check out your result like this...

```{r}
#| eval: false
View(compiled_data)
```


# Importing Data

```{r}
dir_ls(path="data/raw")
csv_files<-dir_ls(path="data/raw")
read_csv(csv_files)
```
