---
title: "Data Analysis"
author: 
  - name: Beau Coomer
  - name: Emmy Delekta
format: 
  html:
    theme: minty
    toc: true
    toc-depth: 2
    highlight-style: github
    code-tools: true
    fig-align: center
editor: source
editor_options: 
  chunk_output_type: console
---

# Importing Data

You always want to load your packages at the top of your script/ document.

```{r}
#| warning: false

library(tidyverse)
library(readr)
```

You can have functions from R packages (like `dplyr`), but you can also write R functions yourself. Load the custom function [clean_keyence_csv](custom_functions/clean_keyence_csv.R)...

```{r}
# Load clean_keyence_csv()
source("custom_functions/clean_keyence_csv.R") 
```

## Dataset 1: Egg Measures

Use the custom function that was imported like this...

1.  Get your file names

```{r}
file_paths <- list.files(path = "data/raw", pattern = "\\.csv$", full.names = TRUE)
```

2.  Process it all into one tibble (similar to data.frame)

```{r}
  compiled_data <- suppressWarnings(
    file_paths %>%
      set_names(basename(.)) %>%
      map(clean_keyence_csv) %>%
      list_rbind(names_to = "source_file")
  )
```

Then, you can check out your result like this...

```{r}
#| eval: false
View(compiled_data)
```

## Dataset 2: Body Measures

```{r}
body_measures <- read.csv("data/crayfish_body_measurements.csv")
```

# Cleaning Data

First, subset the `compiled_data` object to drop irrelevant columns. The `!` tells `select()` to keep everything *except* the columns listed. 

```{r}
clean_csvs<-compiled_data %>%
  select(!c(Measure,date_saved))

call("clean_csvs") # What is call() for??? -Emmy
```

It's best practice to have all names in R to be easily type-able. We can accomplish this by...

* No punctuation at the beginning or end of a string
* Using all lowercase

You can do a direct rename of columns in R by `rename(new_name = old_name)`. If needing to change a bunch of column names, use `rename_with()`.

```{r}
clean_csvs<-clean_csvs %>%
  rename(egg_no=No.) %>%
  rename_with(tolower)
```

All of the important identifying information is kept in the file names. 

```{r}
#| eval: false
# One way to split file names into usable chunks...
str_split_fixed(clean_csvs$source_file,"\\-",4)
```

```{r}
# Better way to split the file names into usable chunks + keep the info
# -- in the data frame
tmp<-clean_csvs %>%
  separate_wider_delim(cols = source_file,delim = "-",names = c("species","treatment","ind_num",NA))
```

```{r}
#| eval: false
View(tmp)
```

Sometimes R will yell at if the data type isn't appropraite. You can check the data type of anything using `class()`.

```{r}
# Works on objects
class(tmp)

# And it can also work on parts of the object
class(tmp$result)
```


If the class isn't what it should be, you can fix it on the fly using one of the many `as.___()` functions. Just know that sometimes, it is not going to work and it may not always be appropriate!

```{r}
tmp<-tmp %>%
  mutate(result=as.numeric(result))

# "$result" is ambiguous and could be misleading, so rename for clarity
tmp<-tmp %>%
rename(egg_diameter_mm=result)
```

```{r}
# Make object to use for summary stats
df_summaries <- unite(tmp,treatment,ind_num,col = "treatment_ind_number",sep = "_")

# Make data that keeps those columns separate
df_analysis <- tmp
```


# Summary Statistics

## Average egg diameter

```{r}
egg_diameter_avgs<-df_summaries %>%
  group_by(treatment_ind_number) %>%
  summarize(avg = mean(egg_diameter_mm))

egg_diameter_avgs
```

## Eggs per individual across treatments
```{r}
ind_egg_counts<-df_summaries %>%
  group_by(treatment_ind_number) %>%
  summarize(egg_count = n())

ind_egg_counts
```

# Analysis

## Experimental Design

**10** *Procambarus clarkii* individuals, each with eggs split across **4** preservation treatments (EtOH80, EtOH95, Iso70, Formalin10). There are not necessarily the same number of eggs placed in each preservation treatment from a given individual.

You can check to make sure the data we're using matches what it should be...

```{r}
#| echo: false

# Number of individuals
length(unique(df_analysis$ind_num))

# List of treatments
unique(df_analysis$treatment)
```

The variables of this study are...

| Type | Variable(s)  |
| ---- | ------------ |
| Independent | Treatment, Individual* | 
| Dependent | Egg size (diameter in mm) | 
| Controlled | Species, Individual* | 

The [assumptions](https://www.statology.org/parametric-tests-assumptions/) for most parametric tests are:

1. Normality
2. Equal variance
3. Independence - data is sampled randomly and independently (not biased)
4. No outliers (no extremes)

Our experimental design violates the assumption of independence --- there are multiple egg measurements per individual. We have **repeated measures**, which, if ignored, undermines the statistical sanctity of the results (i.e., your interpretation would not be valid if you just ignore it). 

To account for repeated measures, you can explicitly model the correlation between that of the dependent variable (egg size (continuous)) and the source of the repeated measure (individual) using a **linear mixed-effect model** or **repeated-measures ANOVA**. 

Since the eggs from each individual were spread across the four treatments, the experimental design is **crossed**. There were not an equal number of eggs per individual/ per treatment, it is also **unbalanced**.

## Linear Mixed-Effect Model (LME)

The R package `{lme4}` is widely used for linear mixed effect modelling. The base R function `lm()` is for linear models, but it can't handle experimental designs with random factors, so that's why we need `{lme4}`.

There are a couple considerations in the `lmer()` function that we'll want to overload. The R package `{lmerTest}` does this, so we'll load it too.

```{r}
#| message: false
#| warning: false
library(lme4)
library(lmerTest)
```

::: {.callout-tip collapse="true" title="Learn more about these models!"}
You can learn more about linear models in Chapter 18 of *The Analysis of Biological Data* by Whitlock and Schluter (3rd Edition, ISBN-10: 131922623X) --- [(R Code for Examples)](https://whitlockschluter3e.zoology.ubc.ca/RExamples/Rcode_Chapter_18.html). 

Dynamic Ecology has a interesting blog post on effects... [Is it a fixed or random effect?](https://dynamicecology.wordpress.com/2015/11/04/is-it-a-fixed-or-random-effect/)

There's also a great chapter on [**Linear Mixed Models** from a Washington U course](https://courses.washington.edu/psy524a/_book/linear-mixed-models.html).
:::

For the model, the user defines the response, fixed effect, and random effect(s). 

$$\mathrm{Egg Diameter}_{ij} = \mathrm{Treatment}_i + (1| \mathrm{Individual}_j)$$

Since every individual has eggs across all four treatments, the model can cleanly separate the effect of the treatment from the unique characters of the individual. `lmer` uses **maximum likelihood** (ML) estimates that naturally weight individuals with more eggs heavier for calculating individual traits while still accurately calculating the overall effect of treatment. The model also pulls/ "shrinks" extreme averages toward the global mean, which prevents a single individual with few eggs from potentially skewing the result.

```{r}
df_analysis$ind_num <- as.factor(df_analysis$ind_num)
df_analysis$treatment <- as.factor(df_analysis$treatment)
```

```{r}
# Create the model
mod_full <- lmer(egg_diameter_mm ~ treatment + (1|ind_num) + (1|treatment:ind_num), data = df_analysis)

# View the result
mod_full
```

### Overall Treatment Effect

Passing the output of `lmer()` to `anova()` gives you the ANOVA table.

```{r}
anova(mod_full)
```

P < 0.05, so **treatment significantly affects the egg diameter**, but it does not give more information as to *which* treatments differ. We'll need to do pairwise comparisons to figure that out.

### Pairwise Comparisons

The treatment is significant, so we can use the R package `{emmeans}` (estimated marginal means). 

```{r}
#| message: false
#| warning: false
library(emmeans)
```

```{r}
#| message: false
emm <- emmeans(mod_full, ~ treatment)
```

```{r}
pairs(emm, adjust = "tukey")
```

::: {.callout-note}
Look at `p.value` in the output. Any value under 0.05 means the two treatments (in `contrast`) are statistically different.
:::

Mean egg diameter differed in all preservative types.

```{r}
summary(emm)
```

Linear mixed models assume the residuals are normally distributed and have equal variance. The study design is unbalanced, so verifying assumptions is even more important.

```{r}
# Check variance
plot(mod_full)

# Visual check for normality of residuals
qqnorm(residuals(mod_full), main = "Q-Q Plot of Residuals")
qqline(residuals(mod_full), col = "red", lwd = 2)

# Statistical check for normality of residuals
shapiro.test(residuals(mod_full))
```

The residuals are falling pretty darn far from the line, especially on the lower end. Since we are dealing with linear measurement data, this kind of normality violation is kind of expected and we can try to transform the data and re-run the model...

```{r}
df_analysis$log_egg_diameter_mm <- log(df_analysis$egg_diameter_mm) + 1
```

```{r}
mod_log <- lmer(log_egg_diameter_mm ~ treatment + (1|ind_num) + (1|treatment:ind_num), data = df_analysis)
```

Now we can check the assumptions again before moving forward...

```{r}
# Check variance
plot(mod_log)

# Check normality of residuals
qqnorm(residuals(mod_log), main = "Q-Q Plot of Residuals (log-transformation)")
qqline(residuals(mod_log), col = "red", lwd = 2)
```

Still, normality is not fantastic, but it's slightly better.

Another assumption is equal variances. We can check this visually, or with a statistical test.

```{r}
#| include: false
#| eval: true
library(car)
```

```{r}
# Visual Check: Points should be random cloud/ smear
plot(mod_log, type = c("p", "smooth"),
     main = "Residuals vs Fitted Values",
     xlab = "Fitted Values", ylab = "Residuals")

# Levene's Test: p > 0.05 means variance is equal across groups (treatments)
leveneTest(residuals(mod_log) ~ df_analysis$treatment)
```

Variances are not equal, so we can either proceed with a huge amount of caution in interpretation or start to use conservative approaches. 

```{r}
anova(mod_log)
mod_log.out <- summary(mod_log)

total_variance <- sum(as_tibble(mod_log.out$varcor)$vcov)

ind_variance <- as_tibble(mod_log.out$varcor) %>%
  filter(grp == "ind_num") %>%
  pull(vcov)

# Interclass Correlation Coefficient (ICC)
ind_variance / total_variance
```

48.8% of all variation in egg diameter is simply due to differences between individuals...

```{r}
emm_log <- emmeans(mod_log, ~ treatment)
pairs(emm_log, adjust = "tukey")
```

All treatments are remain different from one another... and we have the issue of unequal variances.

## Going Further, the Conservative Route

Having all the treatments be statistically significant from one another is just down right weird. We know that the residuals of the model were skewed, even after the egg diameters were log transformed. Since the normality assumption is violated, the p-values are probably artificially deflated. We need to know if our treatment differences are real, so we can use a **generalized linear mixed model** (GLMM) with a link function.

```{r}
#| eval: true
#| include: true
#| warning: false
#| message: false
robust_mod <- glmer(egg_diameter_mm ~ treatment + (1|ind_num), family = Gamma(link = "log"), 
                    data = df_analysis)
library(car)
library(DHARMa)
Anova(robust_mod, type = "III")

residuals_sim <- simulateResiduals(fittedModel = robust_mod)
plot(residuals_sim)
```

::: {.callout-note}
The **Gamma** distribution called from `family =` tells the function to expect the skewed data. It adjusts the standard errors, so the pairwise comparisons might be more realistic. 

Read more about link functions on R-bloggers: [Generalized Linear Models understanding the link function](https://www.r-bloggers.com/2018/10/generalized-linear-models-understanding-the-link-function/)
:::

## Data Cleaning Method...

It seems like the positive skew could be caused by a handful of outliers that could be damaged eggs, measurement errors, or potentially unviable runts. These outliers can be pulled out of the data set systematically using Cook's distance...
```{r}
# install.packages("influence.ME")
library(influence.ME)
```

```{r}
#| eval: false
# Get influence of each observation
infl <- influence(mod_full, obs = TRUE) 
```

:::{.callout-warning}
Running `influence()` takes a looong time. To save time, save it as an RData object (.RDS) with `saveRDS()` and you can use it in the future by loading it with `readRDS()`.
:::

```{r}
#| eval: false
saveRDS(infl, file = "data/rds_objects/infl.rds")
```

```{r}
#| label: "Load infl.rds"
#| eval: true
infl <- readRDS("data/rds_objects/infl.rds")
```

```{r}
cooks <- cooks.distance(infl)

# View observations with high leverage
# -- Cook's distance > 4/n is a standard rule
high_influences <- which(cooks > (4 / nrow(df_analysis)))

# Subset data to get rid of high influences...
df_sub <- df_analysis %>%
  slice(-high_influences)

# Fit model without the outliers
mod_sub <- lmer(egg_diameter_mm ~ treatment + (1|ind_num), data = df_sub)

# Visual check of assumptions
library(performance)
check_model(mod_sub)
# -- Normality issue fixed, but variance looks slightly off.

# Statistical test of assumptions
shapiro.test(residuals(mod_sub))
leveneTest(residuals(mod_sub) ~ df_sub$treatment)
```

Normality problem is fixed with removing outliers, but there is an issue of variance. Linear mixed models are vulnerable to unequal variances. It is a good idea to move forward trying a more robust approach.

## Robust Model: `nmle` {#sec-robust-model}

```{r}
#| warning: false
#| message: false
# 1. Load the packages
library(nlme)
library(car)      # For Type III Anova
library(emmeans)  # For post-hoc pairwise comparisons
```

```{r}
# 2. Fit the model using Generalized Least Squares via lme
# random = ~ 1 | ind_num/treatment builds the nested random interaction term
# weights = varIdent(form = ~ 1 | treatment) models unique variances per group
nlme_model <- lme(fixed = egg_diameter_mm ~ treatment, 
                  random = ~ 1 | ind_num/treatment, 
                  weights = varIdent(form = ~ 1 | treatment), 
                  data = df_sub,
                  method = "REML")
```

```{r}
# 3. View the model summary (shows the estimated standard deviations per treatment)
summary(nlme_model)
```

```{r}
# 4. Generate the Type III ANOVA table 
# "marginal" in nlme is equivalent to Type III Sum of Squares
anova(nlme_model, type = "marginal")
# Anova(nlme_model, type = "III")
```

```{r}
# 5. Conduct post-hoc Tukey adjustments to see which treatments differ
treatment_means <- emmeans(nlme_model, ~ treatment)
pairwise_comps <- pairs(treatment_means, adjust = "tukey")
pairwise_comps
```

# Visualization

The best (statistically sound model) is the [robust model](#sec-robust-model) from `{nmle}`.

* Accomodates randomized block design with subsampling and unequal design
* Potential outliers were removed using a Cook's distance threshold
* Used `varIdent` function to address heteroscedasticity
* Individuals and treatment-by-individual interaction were modeled as nested random effects.
* Significant effect of treatment on egg diameter (Type III ANOVA)
* Post-hoc pairwise comparisons with Tukey adjustments
  - Iso70 had larger egg diameters than all other groups
  - No difference between EtOH80 and EtOH95
  - Formalin10 resulted in significantly smaller egg sizes than all other treatments.

## Step 1: Prepare the estimated means data
You need to extract the clean statistical means from the model. We already have `treatment_means`, but it needs to be turned into a form that can be used by our graphing method.

```{r}
class(treatment_means)

df_emmeans <- as.data.frame(treatment_means)
head(df_emmeans)
```

## Step 2: Make the ggplot object 

I want the treatment categories on the plot to be in order from smallest to largest. 
```{r}
df_sub %>%
  dplyr::select(treatment, egg_diameter_mm) %>%
  group_by(treatment) %>%
  summarize(mean = mean(egg_diameter_mm))
```

```{r}
library(ggplot2)

df_sub %>%
  # PREP: Put categories in order small to large
  mutate(treatment = fct_reorder(treatment, egg_diameter_mm, .fun = 'mean')) %>%
ggplot(aes(x = treatment, y = egg_diameter_mm)) +
  
  # Layer 1: Individual eggs colored by treatment
  geom_jitter(aes(color = as.factor(treatment)), alpha = 0.2, width = 0.2, size = 1) +
  # scale_color_manual(values = c("Formalin10" = "#ffb000", "Iso70" = "#dc267f", "EtOH95" = "#785ef0", "EtOH80" = "#fe6100")) + 
    scale_color_manual(values = c("Formalin10" = "#FF837A", "Iso70" = "#82D69A", "EtOH95" = "#73B2FF", "EtOH80" = "#FFC66D")) + 

  # Layer 2: Boxplots
  geom_boxplot(alpha = 0.5, 
               outlier.shape = NA, 
               fill = "white", color = "black", width = 0.5, staple.linewidth = 0.5, staple.color = "black", 
               staple.linetype = "solid", staplewidth = 0.5) +
  
  # Layer 3: Estimated marginal means
  geom_point(data = df_emmeans, aes(x = treatment, y = emmean),
             color = "red", shape = 18, size = 3) +
  
  # Layer 4: Labels & theme customization
  labs(
    title = "Effect of Preservation Treatment on Egg Diameter",
    subtitle = "Red diamonds represent model-estimated marginal means",
    x = "Treatment Group",
    y = "Egg Diameter (mm)",
    color = "Individual ID (Block)"
  ) + 
  theme_classic(base_size = 14) + 
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "gray30"),
    legend.position = "none",
    axis.title = element_text(face = "bold"),
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold")
  ) +
  ylim(1.5,2.5)
```



```{r}
#| eval: true
#| include: true

# THIS IS UNFINISHED
library(multcomp)
## Generate significance letters ##
treatment_means
# Convert emmeans object to general linear hypothesis object (glht)
letters_data <- cld(treatment_means, Letters = letters, adjust = "none")
# NOTE: Setting adjust = "none" stops emmeans from warning about the CIs and it
# -- should still pas the Tukey-adjusted pairwise test behind the scenes
letters_df <- as.data.frame(letters_data)
# Clean whitespace in letter column...
letters_df$.group <- trimws(letters_df$.group)
```

```{r}
#| eval: true
#| include: true
# Calculate each group's max.
# -- This makes it so you can plot a "floating" letter instead of static.
group_max <- df_sub %>%
  group_by(treatment) %>%
  summarize(max_y = max(egg_diameter_mm, na.rm = TRUE))

# Append the "heights" to the df
letters_df <- left_join(letters_df, group_max, by = "treatment")
```


```{r}
#| eval: true
#| include: true
library(ggplot2)
# library(ggokabeito)

# 5. Plot raw data boxplots with overlaid Tukey letters
fig1 <- df_sub %>%
  # PREP: Put categories in order small to large
  mutate(treatment = fct_reorder(treatment, egg_diameter_mm, .fun = 'mean')) %>%
ggplot(aes(x = treatment, y = egg_diameter_mm)) +
  
geom_jitter(aes(color = as.factor(treatment)), alpha = 0.5, width = 0.2, size = 1) +
 # scale_color_manual(values = c("Formalin10" = "#ffb000", "Iso70" = "#dc267f", "EtOH95" = "#785ef0", "EtOH80" = "#fe6100")) +
  scale_color_manual(values = c("Formalin10" = "#FF837A", "Iso70" = "#82D69A", "EtOH95" = "#73B2FF", "EtOH80" = "#FFC66D")) + 

# ggplot(df_sub, aes(x = treatment, y = egg_diameter_mm)) +
  geom_boxplot(alpha = 0.2, 
               outlier.shape = NA, 
               fill = "white", color = "black", width = 0.5, staple.linewidth = 0.5, staple.color = "black", 
               staple.linetype = "solid", staplewidth = 0.5) +
# ANNOTATION LAYER: Places each letter exactly above its group's highest point
   geom_text(data = letters_df, 
             aes(x = treatment, 
                 y = max_y + 0.05, 
                 label = .group),
             vjust = 0, 
             fontface = "bold", 
             size = 5) +
    # Layer 3: Estimated marginal means
  geom_point(data = df_emmeans, aes(x = treatment, y = emmean),
             color = "black", shape = 21, size = 2.75, fill = "lightgray", stroke = 0.2)  +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.12))) + 

  # Layer 4: Labels & theme customization
  labs(
    #title = "Effect of Preservation Treatment on Egg Diameter",
    #subtitle = "Red diamonds represent model-estimated marginal means",
    x = "Treatment Group",
    y = "Egg Diameter (mm)",
    # color = "Individual ID (Block)"
  ) + 
  scale_fill_okabe_ito(name = "Treatment", order = 3:6) +
  theme_classic(base_size = 14) + 
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "gray30"),
    legend.position = "none",
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    axis.title = element_text(face = "bold"),
  ) +
  scale_x_discrete(labels = c("Formalin10" = "Formalin (10%)", "EtOH95" = "Ethanol (95%)", "EtOH80" = "Ethanol (80%)", "Iso70" = "Isopropanol (70%)"))

fig1
ggsave("fig1.svg", plot = fig1, width = 6, height = 4, dpi = 300)
fig1
```


:::{.callout-note}
The R package `{ggokabeito}` uses the well-known accessible palette 'Okabe-Ito'. By default, it uses colors in the regular order of the palette, but you can explicitly tell R which colors to use if you want to.

To preview what hex codes would look like, you can use `show_col()` from the R package `{scales}`.
:::